FictionEro2 - Affective Impact of Fiction

### 2025-08-04 study2/analysis/2_emotion.qmd
Update 2_emotion.qmd

marginaleffects::avg_slopes(m_arousal, variables=c("Condition", "StimuliType"), by="Gender") |>
  results_table()
```

@@

```{=html}
<!-- 1.For Males, giving feedback indicating that AI-generated images were less attractive, influenced arousal ratings for images labeled as AI-generated. Specifically, males who reported finding AI-generated images as less arousing had lower arousal ratings for those images.

<!-- 2. For Males, giving feedback indicating that AI-generated images were more attractive significantly influenced arousal ratings for images labeled as AI-generated. Specifically, males who reported finding AI-generated images more attractive had higher arousal ratings for these images.  -->
```

@@

display(moderator_results)
```

@@

```{r Enticement Model}
m_enticement <- glmmTMB::glmmTMB(Enticing ~ Gender / StimuliType/ Condition + 
                                   (Condition | Participant) + (1|Stimulus),
                      data=df, family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

@@

marginaleffects::avg_slopes(m_enticement_st, variables="Condition", by="Gender") |>
  results_table()

@@

marginaleffects::avg_slopes(m_valence_st, variables="Condition", by="Gender") |>
  results_table()

@@

marginaleffects::avg_slopes(m_emotionality_st, variables="Condition", by="Gender", re.form=NA) |>
  results_table()
```

@@

-   Gender is a major predictor of arousal, enticement and valence, with females generally reporting lower ratings than males overall.
-   Women prefer couple images (i.e.,significantly higher arousal, enticement and valence ratings); men prefer individual images (i.e.,significantly lower arousal, enticement and valence ratings).
-   AI-generation reduces arousal for men, especially for individual images. AI-generation has no significant effect on arousal for women, whether images are of individuals or couples.

@@

-   For women, positive attitudes towards AI moderated arousal and enticement, resulting in more favorable evaluations of AI-generated images. Negative attitudes moderated valence ratings, leading to less favorable evaluations.
-   Negative attitudes towards AI moderated valence ratings for both genders
-   Reporting having higher knowledge about AI moderated valence ratings for men
-   Sexual activity moderated enticement ratings for females. Specifically, females who reported having sexual activity within the 3 previous days tended to rate AI images higher in enticement.
-   Men’s perceptions of the attractiveness of AI-generated images moderated both arousal and valence ratings. More attractive AI images were rated as more arousing and valent.
-   Female's feedback that no images were attractive moderated arousal ratings.

@@

stars1b <- data.frame(label=c("*","**", "*","*"),
                    Gender=fct_relevel(c("Male", "Male", "Male","Male"), "Male"),
                    Outcome=c("Arousal", "Enticement", "Enticement","Valence"),
                    x=c(1, 1, 2, 1),
                    y=c(0.60, 0.65, 0.56, 0.68))

p1 <- pred1 |>
  mutate(Gender = fct_relevel(Gender, "Male", "Female")) |> 
  ggplot(aes(x=StimuliType, y=Proportion)) +
  geom_hline(yintercept=0, color="darkgrey") +
  geom_line(aes(group=Condition, color=Condition), position = position_dodge(width=0.2), alpha=0.8, linewidth=0.4, key_glyph = "path") +
  geom_pointrange(aes(ymin = CI_low, ymax=CI_high, color=Condition), position = position_dodge(width=0.2), size=0.3) +
  geom_text(data=stars1b, aes(x=x, y=y, label=label), hjust=0.5, color="#424242", size=6) +
  geom_rect(data=data.frame(StimuliType = "Individual", Proportion=0.1, Gender=as.factor("Male"), Outcome="Arousal"),
            aes(xmin=0.8, xmax=1.2, ymin=0.35, ymax=0.7), color="#8BC34A", alpha=0, ) +
  geom_rect(data=data.frame(StimuliType = "Individual",Proportion=0.1, Gender=as.factor("Female"), Outcome="Arousal"),
            aes(xmin=0.8, xmax=1.2, ymin=0.2, ymax=0.5), color="#8BC34A", alpha=0) +
  facet_grid(Outcome~Gender, switch="y", scales ="free") +
  scale_y_continuous(limits=c(0, 1), labels=scales::percent) +
  scale_color_manual(values=c("AI-Generated"="#2196F3", "Photograph"="#F44336")) +
  theme_minimal() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 8),
        strip.placement = "outside",
        strip.background.x = element_rect(fill=c("lightgrey"), color=NA),
        strip.text.x = element_text(size = 10),
        strip.text.y = element_text(size = 10),
        axis.text.x = element_text(angle=25, hjust=1, vjust=1.1, size=9, color="black"),
        legend.text = element_text(size = 10),
        legend.position = ) +
  labs(color="Images presented as:")

@@

```{r}
m1 <- glmmTMB::glmmTMB(Arousal ~ Gender / Condition + (Condition | Participant) + (1|Stimulus),data = df, 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))
m2 <- glmmTMB::glmmTMB(Arousal ~ Gender / Condition + (Condition | Participant) + (1|Stimulus),
                      data=filter(df, ConditionBelief == "True"), 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))
m3 <- glmmTMB::glmmTMB(Arousal ~ Gender / Condition  + (Condition | Participant) + (1|Stimulus),
                      data=filter(df, ConditionBelief == "True", Feedback_NoFacesAttractive=="True"), 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))
m4 <- glmmTMB::glmmTMB(Arousal ~ Gender / Condition  + (Condition | Participant) + (1|Stimulus),
                      data=filter(df, ConditionBelief == "True", Feedback_AIMoreAttractive=="True"), 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

stars2 <- data.frame(label=c("**", "***", "*", "*", "*"),
                    Gender=fct_relevel(c("Male", "Male", "Male", "Female", "Female")),
                    Type=c("All trials", "Believed trials" ,"'AI was more attractive'", "Believed trials", "'No faces were attractive'"),
                    y=c(0.52, 0.56, 0.66, 0.47, 0.39))

pred2 <- rbind(
  estimate_relation(m1) |> 
    mutate(Type = "All trials"),
  estimate_relation(m2) |> 
    mutate(Type = "Believed trials"),
  estimate_relation(m3) |> 
    mutate(Type = "'No faces were attractive'"),
  estimate_relation(m4) |> 
    mutate(Type = "'AI was more attractive'"))

p2 <- pred2 |> 
  mutate(Type = fct_relevel(Type, "All trials", "Believed trials", "'No faces were attractive'", "'AI was more attractive'")) |> 
  ggplot(aes(x=Type, y=Predicted)) +
  geom_hline(yintercept=0, color="darkgrey") +
  geom_line(aes(group=Condition, color=Condition), position = position_dodge(width=0.2), alpha=0.8, linewidth=0.1) +
  geom_bar(stat="identity", aes(fill=Condition), position = position_dodge(width=0.2), width=0.2, alpha=0.6) +
  geom_pointrange(aes(ymin = CI_low, ymax=CI_high, color=Condition), position = position_dodge(width=0.2), size=0.2) +
  geom_rect(data=data.frame(Type="All trials", Predicted=0.1, Gender=as.factor("Male")),
            aes(xmin=0.7, xmax=1.3, ymin=0.35, ymax=0.55), color="#8BC34A", alpha=0) +
  geom_rect(data=data.frame(Type="All trials", Predicted=0.1, Gender=as.factor("Female")),
            aes(xmin=0.7, xmax=1.3, ymin=0.30, ymax=0.48), color="#8BC34A", alpha=0) +
  geom_text(data=stars2, aes(y=y, label=label), hjust=0.5, size=3) +
  facet_grid(~Gender) +
  scale_y_continuous(labels=scales::percent) +
  scale_color_manual(values=c("AI-Generated"="#2196F3", "Photograph"="#F44336")) +
  scale_fill_manual(values=c("AI-Generated"="#2196F3", "Photograph"="#F44336")) +
  labs(y="Arousal",
       tag = "B") +
  theme_minimal() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle=25, hjust=1, vjust=1.1, size=9),
        axis.text.y = element_text(size = 8),
        axis.title.y = element_text(size = 10),
        strip.placement = "outside",
        strip.background.x = element_rect(fill=c("lightgrey"), color=NA),
        strip.text.x = element_text(size = 10),
        legend.position = "none")

>>>>

marginaleffects::avg_slopes(m_arousal, variables="Condition", by="Gender") |>
  results_table()
```

@@

<!-- 1.For Males, giving feedback indicating that AI-generated images were less attractive, influenced arousal ratings for images labeled as AI-generated. Specifically, males who reported finding AI-generated images as less arousing had lower arousal ratings for those images.

<!-- 2. For Males, giving feedback indicating that AI-generated images were more attractive significantly influenced arousal ratings for images labeled as AI-generated. Specifically, males who reported finding AI-generated images more attractive had higher arousal ratings for these images.  -->

@@

display(moderator_results)

@@

```{r Enticement Model}
m_enticement <- glmmTMB::glmmTMB(Enticing ~ Gender / Condition + 
                                   (Condition | Participant) + (1|Stimulus),
                      data=df, family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

@@

marginaleffects::avg_slopes(m_enticement_st, variables=c("Condition", "StimuliType"), by="Gender") |>
  results_table()

@@

marginaleffects::avg_slopes(m_valence_st, variables=c("Condition", "StimuliType"), by="Gender") |>
  results_table()

@@

marginaleffects::avg_slopes(m_emotionality_st, variables=c("Condition", "StimuliType"), by="Gender", re.form=NA) |>
  results_table()
```

@@

-   Gender is a major predictor of arousal, enticement and valence, with females generally reporting lower ratings than males overall but not significantly.
- Men significantly rate AI images as less arousing, enticing and valent than images presented as real photos.
- Women rate images presented as AI-generated lower in valence than images presented as real photos.

-  Women prefer couple images (i.e.,significantly higher arousal, enticement and valence ratings); men prefer individual images (i.e.,significantly lower arousal, enticement and valence ratings).
- AI-generation reduces arousal for men, especially for individual images. AI-generation has no significant effect on arousal for women, whether images are of individuals or couples.

@@

<!-- -   For women, positive attitudes towards AI moderated arousal and enticement, resulting in more favorable evaluations of AI-generated images. Negative attitudes moderated valence ratings, leading to less favorable evaluations. -->
<!-- -   Negative attitudes towards AI moderated arousal, valence ratings for both genders -->
<!-- -   Reporting having higher knowledge about AI moderated arousal, valence ratings for men -->
<!-- <!-- -   Sexual activity moderated enticement ratings for females. Specifically, females who reported having sexual activity within the 3 previous days tended to rate AI images higher in enticement. --> -->
<!-- -   Men’s perceptions of the attractiveness of AI-generated images moderated both arousal and valence ratings. More attractive AI images were rated as more arousing and valent, less attractive. -->
<!-- -   Female's feedback that no images were attractive moderated arousal ratings. -->

@@

# stars1b <- data.frame(label=c("*","**", "*","*"),
#                     Gender=fct_relevel(c("Male", "Male", "Male","Male"), "Male"),
#                     Outcome=c("Arousal", "Enticement", "Enticement","Valence"),
#                     x=c(1, 1, 2, 1),
#                     y=c(0.60, 0.65, 0.56, 0.68))

p1 <- pred1 |>
  mutate(Gender = fct_relevel(Gender, "Male", "Female")) |> 
  ggplot(aes(x=Outcome, y=Proportion)) +
  geom_hline(yintercept=0, color="darkgrey") +
  geom_line(aes(group=Condition, color=Condition), position = position_dodge(width=0.2), alpha=0.8, linewidth=0.4, key_glyph = "path") +
  geom_pointrange(aes(ymin = CI_low, ymax=CI_high, color=Condition), position = position_dodge(width=0.2), size=0.3) +
  # geom_text(data=stars1a, aes(x=x, y=y, label=label), hjust=0.5, color="#424242", size=6) +
  # geom_rect(data=data.frame(StimuliType = "Individual", Proportion=0.1, Gender=as.factor("Male"), Outcome="Arousal"),
  #           aes(xmin=0.8, xmax=1.2, ymin=0.35, ymax=0.7), color="#8BC34A", alpha=0, ) +
  # geom_rect(data=data.frame(StimuliType = "Individual",Proportion=0.1, Gender=as.factor("Female"), Outcome="Arousal"),
            # aes(xmin=0.8, xmax=1.2, ymin=0.2, ymax=0.5), color="#8BC34A", alpha=0) +
  facet_grid(~Gender, switch="y", scales ="free") +
  scale_y_continuous(limits=c(0, 1), labels=scales::percent) +
  scale_color_manual(values=c("AI-Generated"="#2196F3", "Photograph"="#F44336")) +
  theme_minimal() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 8),
        strip.placement = "outside",
        strip.background.x = element_rect(fill=c("lightgrey"), color=NA),
        strip.text.x = element_text(size = 10),
        strip.text.y = element_text(size = 10),
        axis.text.x = element_text(angle=25, hjust=1, vjust=1.1, size=9, color="black"),
        legend.text = element_text(size = 10),
        legend.position = ) +
  labs(color="Images presented as:")

@@

```{r}
m1 <- glmmTMB::glmmTMB(Arousal ~ Gender / Condition + (Condition | Participant) + (1|Stimulus),data = df, 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))
m2 <- glmmTMB::glmmTMB(Arousal ~ Gender / Condition + (Condition | Participant) + (1|Stimulus),
                      data=filter(df, ConditionBelief == "True"), 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))
m3 <- glmmTMB::glmmTMB(Arousal ~ Gender / Condition  + (Condition | Participant) + (1|Stimulus),
                      data=filter(df, ConditionBelief == "True", Feedback_AILessAttractive=="True"), 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))
m4 <- glmmTMB::glmmTMB(Arousal ~ Gender / Condition  + (Condition | Participant) + (1|Stimulus),
                      data=filter(df, ConditionBelief == "True", Feedback_AIMoreAttractive=="True"), 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

stars2 <- data.frame(label=c("**", "***", "*", "***", "*", "*"),
                    Gender=fct_relevel(c("Male", "Male", "Male", "Female", "Male", "Male")),
                    Type=c("All trials", "Believed trials" ,"'AI was more attractive'", "Believed trials", "'AI was less attractive'", "'AI was less attractive'"),
                    y=c(0.52, 0.56, 0.66, 0.47, 0.39, 0.6))

pred2 <- rbind(
  estimate_relation(m1) |> 
    mutate(Type = "All trials"),
  estimate_relation(m2) |> 
    mutate(Type = "Believed trials"),
  estimate_relation(m3) |> 
    mutate(Type = "'AI was less attractive'"),
  estimate_relation(m4) |> 
    mutate(Type = "'AI was more attractive'"))

p2 <- pred2 |> 
  mutate(Type = fct_relevel(Type, "All trials", "Believed trials", "'AI was less attractive'", "'AI was more attractive'")) |> 
  ggplot(aes(x=Type, y=Predicted)) +
  geom_hline(yintercept=0, color="darkgrey") +
  geom_line(aes(group=Condition, color=Condition), position = position_dodge(width=0.2), alpha=0.8, linewidth=0.1) +
  geom_bar(stat="identity", aes(fill=Condition), position = position_dodge(width=0.2), width=0.2, alpha=0.6) +
  geom_pointrange(aes(ymin = CI_low, ymax=CI_high, color=Condition), position = position_dodge(width=0.2), size=0.2) +
  geom_rect(data=data.frame(Type="All trials", Predicted=0.1, Gender=as.factor("Male")),
            aes(xmin=0.7, xmax=1.3, ymin=0.35, ymax=0.55), color="#8BC34A", alpha=0) +
  geom_rect(data=data.frame(Type="All trials", Predicted=0.1, Gender=as.factor("Female")),
            aes(xmin=0.7, xmax=1.3, ymin=0.30, ymax=0.48), color="#8BC34A", alpha=0) +
  geom_text(data=stars2, aes(y=y, label=label), hjust=0.5, size=3) +
  facet_grid(~Gender) +
  scale_y_continuous(labels=scales::percent) +
  scale_color_manual(values=c("AI-Generated"="#2196F3", "Photograph"="#F44336")) +
  scale_fill_manual(values=c("AI-Generated"="#2196F3", "Photograph"="#F44336")) +
  labs(y="Arousal",
       tag = "B") +
  theme_minimal() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle=25, hjust=1, vjust=1.1, size=9),
        axis.text.y = element_text(size = 8),
        axis.title.y = element_text(size = 10),
        strip.placement = "outside",
        strip.background.x = element_rect(fill=c("lightgrey"), color=NA),
        strip.text.x = element_text(size = 10),
        legend.position = "none")
